### 2023-09-15 thesis/MSECE_thesis.tex
Outlier detection via LOF

\begin{table}[h!]
    \centering
    \begin{tabular}{l r r r r}
        \toprule
        \textbf{Corpus} & \textbf{Samples} & \textbf{Mean Sequence Length (\(\pm \sigma\))} & \textbf{Unique Tokens} & \textbf{Tokens} \\
        \midrule
        ATCC            & 29,862           & \(11.90 \pm 7.22\)                             & 2,210                  & 355,476         \\
        ATCO2           & 874              & \(12.46 \pm 5.93\)                             & 855                    & 10,886          \\
        ATCOSIM         & 9,544            & \(11.29 \pm 4.18\)                             & 824                    & 107,759         \\
        ZCU CZ ATC      & 14,435           & \(10.32 \pm 5.50\)                             & 3,106                  & 149,031         \\
        \midrule
        Total           & 54,715           & \(11.39 \pm 7.22\)                             & 5,343                  & 623,152         \\
        \bottomrule
    \end{tabular}
    \label{table:corpora_stats}
    \caption{The number of samples, mean sequence length, number of unique tokens, and total number of tokens by corpus, including the cumulative
        total across all corpora.}
\end{table}

@@

\subsection{Outlier Removal}\label{sec:outlier_removal}
This section analyzes the same corpora as above after outliers in the data have been removed, if any are present. Outlier removal was performed using
the interquartile range (IQR) outlier removal method. IQR is defined as the difference between the medians of the third and first quartiles of the
distribution: \(\mathbf{IQR} = \mathbf{Q3} - \mathbf{Q1}\), where \(\mathbf{Q3}\) and \(\mathbf{Q1}\) are the medians of the third and first quartiles,
respectively. Outliers are then calculated as samples whose sequence lengths lie below \(\mathbf{Q1} - 1.5\mathbf{IQR}\) and above
\(\mathbf{Q3} + 1.5\mathbf{IQR}\).

>>>>

\begin{table}[h!]
    \centering
    \begin{tabular}{l r r r r}
        \toprule
        \textbf{Corpus} & \textbf{Samples} & \textbf{Mean Sequence Length (\(\pm \sigma\))} & \textbf{Unique Tokens} & \textbf{Tokens} \\
        \midrule
        ATCC            & 29,862           & \(11.90 \pm 7.22\)                             & 2,209                  & 355,473         \\
        ATCO2           & 874              & \(12.28 \pm 4.12\)                             & 786                    & 10,733          \\
        ATCOSIM         & 9,544            & \(11.23 \pm 4.12\)                             & 823                    & 107,153         \\
        ZCU CZ ATC      & 14,435           & \(10.05 \pm 5.50\)                             & 2,983                  & 145,107         \\
        \midrule
        Total           & 54,715           & \(11.30 \pm 6.36\)                             & 5,040                  & 618,466         \\
        \bottomrule
    \end{tabular}
    \label{table:corpora_stats}
    \caption{The number of samples, mean sequence length, number of unique tokens, and total number of tokens by corpus, including the cumulative
        total across all corpora.}
\end{table}

@@

\subsection{Outlier Detection and Analysis}\label{sec:outliers}
The Local Outlier Factor (LOF) algorithm~\cite{breunig_lof_2000} was used to detect and remove outliers in the corpora for further analysis. The
sequence lengths of the data samples were used to determine outliers. Since all four corpora are aggregated, LOF analysis was performed on the
aggregated data instead of on the individual corpora.

Sci-Kit Learn's implementation of the LOF
algorithm\footnote{https://scikit-learn.org/stable/modules/generated/sklearn.neighbors.LocalOutlierFactor.html\#sklearn-neighbors-localoutlierfactor}
was used with a value of \(N = 35\) for \(N\) neighbors in the K-Nearest Neighbors algorithm and Minkowski distance for distance metrics/calculations
(in this case, since \(p=2\), Euclidean distance is effectively used for distance calculations). A total of 47 samples were detected as outliers and
removed for analysis and are summarized in table \ref{}

\begin{table}[h!]
    \centering
    \begin{tabular}{l | r}
        \toprule
        \multicolumn{2}{c}{Outlier Properties}   \\
        \midrule
        \textbf{Mean Sequence Length}    & 47.30 \\
        \textbf{Std of Sequence Length}  & 11.82 \\
        \textbf{Minimum Sequence Length} & 38    \\
        \textbf{Maximum Sequence Length} & 73    \\
        \textbf{\# Samples}              & 47    \\
        \bottomrule
    \end{tabular}
    \label{table:outlier_stats}
    \caption{}
\end{table}
